\documentclass[header.tex]{subfiles}
\begin{document}
\begin{center}
\textbf{Authored Publications} \\
\end{center}
%\vspace{0.1cm}

\noindent\textbf{Longitudinal heritability of childhood aggression.} \\
\textbf{Porsch RM}, Middeldorp CM, Cherny SS, Krapohl E, van Beijsterveldt CEM, Loukola A, Korhonen T, Pulkkinen L, Corley R, Rhee S, Kaprio J, Rose RR, Hewitt JK, Sham P, Plomin R, Boomsma DI, Bartels M \\
\textit{American Journal of Medical Genetics, Part B\@: Neuropsychiatric Genetics 2016, 171:697–707.} \\
\vspace{0.5cm}

\noindent \textbf{Sacral agenesis: a pilot whole exome sequencing and copy number study.} \\
\textbf{Porsch RM}, Merello E, De Marco P, Cheng G, Rodriguez L, So M, Sham PC, Tam PK, Capra V, Cherny SS, Garcia-Barcelo M-M, Campbell DD \\
\textit{BMC Medical Genetics 2016, 17:98.} \\
\vspace{0.5cm}

\begin{center}
\textbf{Co-Authored Publications} \\
\end{center}

\noindent\textbf{Polygenic scores via penalized regression on summary statistics.} \\
Mak TSH, \textbf{Porsch RM}, Choi SW, Zhou X, Sham PC \\
\textit{Genetic Epidemiology 2017(February):1–12.} \\
\vspace{0.5cm}

\noindent\textbf{Cost effective assay choice for rare disease study designs.} \\
Campbell DD, \textbf{Porsch RM}, Cherny SS, Capra V, Merello E, De Marco P, Sham PC, Garcia-Barceló M-M \\
\textit{Orphanet Journal of Rare Diseases 2015, 10:10.} \\
\vspace{0.5cm}

\noindent\textbf{Exome-wide association analysis reveals novel coding sequence variants associated with lipid traits in Chinese.} \\
Tang CS, Zhang H, Cheung CYY, Xu M, Ho JCY, Zhou W, Cherny SS, Zhang Y, Holmen O, Au K-W, Yu H, Xu L, Jia J, \textbf{Porsch RM}, Sun L, Xu W, Zheng H, Wong L-Y, Mu Y, Dou J, Fong CHY, Wang S, Hong X, Dong L, Liao Y, Wang J, Lam LSM, Su X, Yan H, Yang M-L, et al. \\
\textit{Nature Communication 2015, 6:10206.} \\
\vspace{0.5cm}

\end{document}
