%! TEX root = /home/robert/Documents/projects/thesis/header.tex

In the previous chapters I have investigated the genetic architecture of aggressive behavior.
I outlined previous studies which examined genetic factors affecting aggression and described general methodological approaches in identifying those.

My first study explored the longitudinal heritability of childhood aggression.
I showed that genetic factors, which influence aggression, are rather stable over age and that quantitative sex differences are minor.
Following, I analysed specific molecular markers which might influence both impulsive aggression and risk taking behavior.
I identified two independent SNPs associated with risk taking, but failed to detect any genome-wide signal for impulsive aggression.
Next, I investigated the relationship of impulsive aggression and risk taking with various psychiatric disorders.
I showed  high genetic correlations between impulsive aggression and depression (MDD and DS).
In addition, application of a Mendelian randomization demonstrated that schizophrenia might causally affect both impulsive aggression and risk taking.
Within my last study I investigated the properties of the KS-Burden test.
This test is designed to explore distributional differences of rare variants between affected and unaffected individuals.
The test showed generally better performance than other rare variant tests when clusters of causal variants were present.
Application of the KS-Burden test to examine the impact of rare variants on aggressive behavior did not result in any genome-wide significant findings, however.

Within this chapter I will discuss my findings in general.
First I am going to contrast heritability estimates from  my twin study and GWAS\@.
Following, I will explore the results of the association studies as well as implications for future studies.
Finally, I will discuss genetic correlations in general, followed by an examination of Mendelian randomizations.

\section{Heritability of Aggression}
\label{sec:heritability_of_aggression}

The missing heritability problem is well-known in genetics.
Thus, it is not surprising to find that heritability estimates from the twin study and those computed on SNPs differ.
Specifically, I showed that heritability of aggressive behavior in children is stable across age and ranged from 50 to 80\%, while the heritability estimate from the conducted GWAS was 5\%. 

First of all it is important to note that these two heritability estimates are not directly comparable.
Specifically, twin estimations were done on children while SNP heritability was derived from mostly middle-aged adults.
However, previous heritability estimations in adult twins were similar estimates to those in children~\cite{Miles1997a}.

The observed differences could also have been caused by the differences in instruments as well as the definition of aggressive behavior.
While both  twin studies made use of a validated psychometric instrument to measure aggressive behavior in children, measurements in the UK Biobank were based on a single question with dichotomous answers choices.
Thus, measurements within the UK Biobank are of less precision and potentially include more noise.
Furthermore, aggressive behavior within the UK Biobank was defined as an impulsive act while the  instruments in the twin studies assumed a more general definition of aggression.
Hence, the  instruments show a considerable degree of diversity in both definition and psychometric properties, potentially affecting heritability estimations.

The GWAS by~\citet{Pappa2016a} also demonstrated great variability across assessed cohorts in terms of SNP heritability of aggressive behavior in children as well ($10-54\%$).
However, it is unclear if these differences are due to genotyping platform, the age of included participants or assessment instrument.
Furthermore, 95\% confidence intervals across cohort estimates are overlapping indicating that differences in heritability are not statistically significant.
Indeed, only the two smaller cohorts of this particular study show relatively high heritability estimates ($h^2=.54, N=2,101; h^2=.46, N=908$), while the largest cohort displayed a similar estimate as the one presented in Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking} ($h^2=.1, N=5,505$).
This suggests that estimated SNP heritability of impulsive aggression within the UK Biobank is similar to that estimated from validated psychometric instruments. 

However, there are a number of potential additional reasons for the observed discrepancy between twin and GWAS heritability estimates as I already described in Section~\ref{sec:heritability_and_genetic_correlation}.
GWAS rarely capture potential epistatic effects, discount the influence of rare variants, and are unable to investigate the influence of regulatory components. 
However, to what extent these additional factors influence the size of the observed discrepancy remains unknown.

Interestingly a study investigating family-based heritability estimations (i.e. using parent-offspring and sibling correlations) showed that models which did take shared environmental factors into account (using SEM), compared to those which did not (Falconer's method), resulted in heritability estimates closer to those seen in genotype data~\citet{Munoz2016a}.
However, this study only looked at diseases such as cardiovascular diseases, various forms of cancers, and diabetes.
Noticeable, the only behavioral  disorder assessed within this study, namely depression, showed the largest missing heritability.
A slightly different suggestion was made by~\citet{Yang2015} who argued that the missing heritability can be explained by taking rare variants into account.
Indeed, in their study rare and common variants were able to explain most of the missing heritability in BMI and height.
While it is unclear to which extent this seems also the case for other phenotypes, most studies to date failed to identify rare variants which were able to account for a considerable proportion of phenotyical variability in a complex trait~\cite{Chabris2015,Wray2011}.
Thus, one can speculate that the simplified instrument used with the GWAS, as well as the different operationalization, as well as definition of aggression has considerable contribution to the discrepancy, but other sources might also be possible.

However, the missing heritability should not distract from other interesting questions such as \acrfull{gxe}.
As shown in Section~\ref{sec:evolutionary_theories}, aggressive behavior has a variety of beneficial and harmful consequences, depending on environmental circumstances. 
For example, while aggressive behavior in social situations can be beneficial to gain social status, it can also be highly penalized by others in the group~\cite{Buss1997}.
Another example is the findings by~\citet{Figueredo1995} which showed that aggressive behavior of husbands towards their wives was profoundly affected by the distance or presence of brothers or powerful fathers.
These findings would suggest considerable environmental influence on the expression of aggressive behavior as well as possible gene-environment interactions (see \textit{MAOA}-environment interactions in Section~\ref{sub:maoa_interactions}).
However, these interactions will not appear in heritability estimates, or the ratio of genetic to total variance in a population.
%, which can also be described as the ratio of genetic and total variance in a population which shares a common environment.
Indeed, the presents of a hidden underlying environmental structure which interacts with genetic factors would push the genetic variance to the denominator.
Thus, reducing the heritability estimation.
Nevertheless, GxE effects might be able to explain a proportion of the total phenotypic variance.

To conclude, the study presented in Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking} represents the first extensive investigation of the genetic architecture of adult human aggression.
The use of over $17,000$ twin pairs not only enabled a robust examination of genetic and environmental factors involved in childhood aggression, but also empowered a detailed analysis of potential sex and age differences (see Chapter~\ref{cha:longHera}).
This analysis demonstrated that sex and age only had a minor effect the estimated heritability.
I further showed that SNP heritability in impulsive aggression is  only 5\%, relatively small.
Next I will discuss my genetic association studies on impulsive aggressive behavior as well as risk taking. 

\section{Genomic Associations in Aggression and Risk Taking}
\label{sec:genomic_associations_in_aggression_and_risk_taking}

\subsection{Common Variants}
\label{sub:common_variants_discussion}

Within Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking} I conducted a GWAS on both risk taking and impulsive aggression.
While I was able to replicate previous findings regarding risk taking~\cite{Day2016}, I was unable to identify any genome-wide signal for impulsive aggression.

Interestingly, genetic variants within \textit{CADM2} which were shown to influence risk taking are also associated with a number of other personality traits~\cite{Boutwell2017},
thus suggesting genetic overlap between risk taking and other personality characteristics.
Indeed, conditional FDR analysis (see Section~\ref{sub:conditional_fdr}) as well as genetic correlations (see Section~\ref{sub:genetic_correlation_ukb_assoc}) between risk taking and neuroticism suggest some overlap.

Furthermore, genetic variants associated with risk taking seem to be in close genomic proximity to those known to influence a proxy phenotype of smoking, namely spirometric measurements (the amount and speed of air a person can inhale/exhale).
Indeed, it is not clear to what extent as well as how risk taking and smoking behaviors are linked.
Previous research has shown that sensation seeking~\cite{Carton1994} and impulsivity~\cite{Glicksohn2007,Mitchell1999} is more prominent in smokers than non-smokers.
Furthermore, a study by~\citet{Ert2013} explored how smokers and non-smokers differ regarding risk taking and found that smokers are more easily tempted by high rewards,
thus suggesting that higher prevalence of risk taking behavior in smokers reflects an impulsive behavior to give into immediate temptations. 
One can therefore speculate that both traits might have shared genetic factors which influence both risk taking and smoking.
Indeed, genetic correlation analysis showed considerable genetic correlation between the two traits (see Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking}).
However, an alternative explanation is that risk taking is causally influencing smoking status,
suggesting that an intervention which is aimed to reduce risk taking behavior might reduce smoking prevalence. 
Nevertheless, it is unclear how such a general personality trait can be experimentally modified to have a long-term health effect.

In contrast to risk taking I was unable to identify any genome-wide associations for impulsive aggression, despite estimating the heritability of aggressive behavior between 50--80\% (see Chapter~\ref{cha:longHera}).
A likely reason for the current failure to identify genome-wide significant variants is that the study presented in Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking} as well as the similar study by~\citet{Pappa2016a} have limited statistical power.
While the study by~\cite{Pappa2016a} on aggression in young children and teenagers is limited to a relatively small sample size ($N=18,988$),
my study, with nearly twice as many subjects, measured aggressive behavior without the help of a validated psychometric instruments and consisted only of one single `yes' or `no' question, unlike~\citet{Pappa2016a}, which made use of well-known instruments to assess aggressive behavior.

Thus, it is obvious that an increase in sample size is necessary in order to boost statistical power.
However, it remains difficult to both obtain high quality phenotype data while increasing sample size to tens of thousands of participants.
The data obtained by~\citet{Pappa2016a}, as well as those used in Chapter~\ref{cha:longHera}, are based on multi-question instruments, commonly used with multiple raters.
While the validity of these tools has been shown multiple times~\cite{Goodman1997,Goodman2001,Achenbach2003}, it is difficult to see how these can be translated to studies with larger sample sizes.
This opens the need to develop instruments which can accurately measure aggressive behavior in a larger samples or validate pre-existing ones.  
For example, the  instrument to measure impulsive aggression in Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking} has not been validated but could easily be explored in the context of other instruments such as CBCL and SDQ\@.
In addition, it is important to note that the distribution of aggression scores of both CBCL and SDQ (see Figure~\ref{fig:hist_aggression}) is skewed.
Therefore, a dichotomous instrument might be able to accurately assess aggressive behavior in larger samples. 

\subsection{Rare Variants}
\label{sub:rare_variants_disccusion}

Rare variants, in respect to impulsive aggression, were analysed in Chapter~\ref{cha:distribuional_differences_of_rare_variants}.
Specifically, I developed a new rare variant association test, called KS-Burden, in order to detect clusters of causal mutations within genomic regions such as genes.
While simulations have shown relatively good performance of KS-Burden compared to other tests, application to impulsive aggression yielded no genome-wide significant findings. 

There are multiple possible reasons for these null findings (see Section~\ref{sec:discussion_ks}).
Most notably, it remains unlikely that rare variants will have a considerable effect on behavioral traits.
Indeed, there is currently no study demonstrating considerable contribution of rare variants in explaining variation in behavioral phenotypes~\cite{Chabris2015}.
However, one could argue that these lack of findings are due to the small sample sizes commonly seen in sequencing-based studies~\cite{Lee2014}. 
In this respect, it is important to emphasis that, in contrast to the association study of common variants, the outlined rare variant association study has considerable statistical power~\cite{Lee2011}.
Specifically, while most studies which aim to explore the impact of rare variants only use a few hundred samples the study presented in Chapter~\ref{cha:distribuional_differences_of_rare_variants} was able to make use of over $30,000$ subjects. 
Thus, this well-powered study could suggest that the observed  null findings imply that rare variants are unlikely to have a major influence on impulsive aggression.

However, it is important to mention that the rare variant association study in Chapter~\ref{cha:distribuional_differences_of_rare_variants} made use of chip array data, a method which is unable to detect private mutations due to the limited number of SNPs adequately tagged.
Indeed, the~\acrfull{exac} has shown in a sample of $60,706$ unrelated individuals that the majority of rare mutations are private mutations (also called singletons), and that a large proportion of these mutations have potential disease implications~\cite{Lek2016,Kobayashi2017}.
However, this does not necessary imply that rare mutations might play an important role in common traits, such as aggression~\cite{Chabris2015}.
Nevertheless, rare, as well as private, mutations could be helpful in explaining extreme expressions of aggressive behavior.
For example, a study by~\cite{Peloso2016} recently showed that extreme phenotype selection results in a greater increase in statistical power in rare variants compared to common variants,
thus suggesting that future studies might look at highly violent criminal offenders in order to identify genes with higher burden of rare variants.

\section{Phenotypic and Genomic Correlations}
\label{sec:phenotypical_and_genomic_correlations}

As described in Chapter~\ref{cha:introduction}, the genetic overlap between aggression and other behavioral phenotypes was unknown.
Chapter~\ref{cha:assocation_study_in_agggressive_behavior_and_risk_taking} as well as Chapter~\ref{cha:psychiatric_corr_mr} explored these genetic relationships.
In particular, I have shown considerable genetic correlations between aggression and risk taking, as well as with neuroticism, smoking, and depression. 
Especially interesting is the unusually high genetic correlation between aggression and depression.

However, as described in Section~\ref{sec:heritability_and_genetic_correlation}, genetic correlations can arise from a multitude of different scenarios.
Thus, it remains unclear if these correlations arise from direct pleiotropic effects, are mediated by other variables, or are caused by other scenarios (see Figure~\ref{fig:genetic_correlation}). 
Indeed, while genetic correlations are able to give some insight into the biological interconnectedness of traits, they suffer some same problems as other correlational analysis. 
For example, the high genetic correlations between depression and aggression could be caused by underlying but unobserved third, fourth, or numerous other phenotypes.

In the past, genetic correlations were only measurable in twin and other genetically-informative studies, or with  raw genotype and phenotype data.
However, with the recent development of LD-score regression, researchers are now able to compute genetic correlations for     a number of phenotypes based on summary statistics alone.
While this had led to an explosion of genetic correlations~\cite{Bulik-Sullivan2015b,Bulik-Sullivan2015a}, it remains unclear how to interpret these genetic correlations regarding disease etiology.

There have been some efforts to understand genetic correlations in a more local sense.
\citet{Shi2016a} (under revision) showed that it is possible to partition genetic correlations across different genetic loci.
The authors showed that the cross trait correlations of 27 traits were mostly driven by only 27 genomic regions, and a number of small genome-wide genetic correlations were highly locally correlated.
While these attempts are important in order to foster our understanding of the shared biological mechanisms across different disorders, they also can be useful in order to understand potential causal relationships across traits.
Indeed,~\citet{Shi2016a} attempted to investigate putative causality by making use of genome-wide significant SNPs within correlation clusters.
For example, the authors demonstrated that the genome-wide correlation between \acrfull{bmi} and \acrfull{tg} can be explained by considering the local correlations around SNPs which were significant in BMI but not TG\@.
In contrast, the local genetic correlation of SNPs significant in TG  but not BMI was not significantly different from $0$.
Hence, suggesting that BMI might have a causal effect on TG\@.
However, while this method is an interesting approach it does not account for potential hidden variables.
Specifically, shared genetic variants might affect a related or sub-phenotype that might be pleiotropic for both BMI and TG\@.
Furthermore, more complex causal relationships, such as those often found in behavioral traits, might be less straightforward than the given example by~\citet{Shi2016a}.
Hence, while local genetic correlations might help to improve our biological understanding it does only partially improve the interpretation.

Another problem in the analysis of genetic correlations is that most summary statistics were obtained from different samples.
This is commonly seen as something positive, given that sample overlap can lead to a bias in the estimation of genetic correlations when using polygenic risk scores. 
However, LD score regression accounts for potential sample overlap~\cite{Bulik-Sullivan2015a} and gives an unbiased correlation estimate.  
Nevertheless, the availability of all relevant phenotype data within a single sample allows one to adjust for potential confounders already at the GWAS state.
This makes the UK Biobank especially useful and further methodological development might be directed at using classical regression methods to further elucidate causal interrelationships among phenotypes variables.

An additional interesting approach to further investigate genetic correlation structures is the use of Mendelian randomization.
I have used this popular approach in Chapter~\ref{cha:psychiatric_corr_mr} to demonstrated suggestive causal effects of schizophrenia on both risk taking and aggression.
However, there are a number of theoretical and practical issues which might lead one to question the validity of MR\@.

\subsection{Mendelian Randomization}
\label{sub:mendelian_randomization_discussion}

\subsubsection{Mendelian Randomization and Randomized Control Trials}
\label{ssub:causality_and_mendelian_randomization}

MR can be seen as similar to classical randomized control trials in which random genetic segregation  assigns subjects to either treatment or control group~\cite{Hingorani2005}.
However, as~\citet{Burgess2016a} pointed out, the aims of randomized control trials and MR are different.
While randomized control trials aim to evaluate the effectiveness of different treatment strategies, the main goal of MR is to assess the magnitude of the causal relationship between an exposure and an outcome. 
Only at a later stage MR aims to use this information to design non-genetic interventions.

This further implies that the effect of any designed intervention will differ from the genetically determined effects estimated by MR\@.
This is due to the length of the intervention (life long versus short term), the mechanism of intervention, as well as the magnitude of the intervention (genetic effects are usually small)~\cite{Evans2015}. 
Hence,~\citet{Vanderweele2015} questioned the use of  numerical estimates in MR and instead suggested testing only for the presences of a causal effect.

Thus, the results presented in Chapter~\ref{cha:psychiatric_corr_mr} have to be viewed in the context of experimental studies as well.
Indeed, as described in Section~\ref{sec:discussion_ukb_psych}, a number of randomized control trials support the notion of a causal relationship between schizophrenia and impulsive aggression as well as risk taking. 
However, it is unclear if other variables might mediate this relationship.

To conclude, while MR has been described in terms of randomized control trials there are some important differences.
However, MR can be useful in mining potential causal connections for future clinical trials~\cite{Evans2015}. 
In this regard it is important to emphasize that large-scale and hypothesis-free approaches to identify potential causal relationships in a large set of phenotypes need to be mindful regarding the underlying assumptions of MR\@.
Indeed, careful testing of assumptions might be difficult when exploring hundreds of potential causal connections,
thus increasing the chance of potential false findings.
However, MR can be a valuable tool in understanding the genetic correlations among multiple interconnected traits. 

\subsubsection{Instrument selection}
\label{ssub:instrument_selection}

A number of studies have been successful in identifying causal relationships between exposures and outcomes.
Most notable studies have shown causal effects of various traits, such has blood pressure, type-2 diabetes, and BMI, on cardiovascular disorders with the help of MR~\cite{Swerdlow,Ference2015,Lieb2013,Voight2012a}.
However, these successes were based on a biological understanding on how the instrument will affect the exposure in question.

This luxury is often not given in many other studies and information about the biological implications of associated variants is unknown.
This scenario is common in behavioral traits, psychiatric disorders, as well as other complex phenotypes.
Thus, some MR studies make use of statistically driven methods to chose SNPs as instrumental variables.
However, as~\citet{Burgess2016a} pointed out, studies which only use a statistically driven choice of instrumental variables and lack clear biological information are not true MR and might better be called `joint-association study' in which the underlying assumptions are only investigated post-hoc.

The lack of biological information has led to the development of new statistical methods to use multiple SNPs as instrumental variables. 
Indeed,~\citet{Burgess2016a} suggested that a specific causal effect of a risk factor on an outcome becomes more plausible if multiple SNPs, which are associated with the risk factor, agree with the effect.
For example,~\citet{Burgess2013} proposed to use allele scores as instrumental variables.
An allele score is either the weighted or unweighted sum of risk-increasing alleles for a given subject.
Indeed, usage of allelic scores from known biological markers facilitated correctly replicating known causal relationships~\cite{Timpson2005,CReactiveProteinCoronaryHeartDiseaseGeneticsCollaborationCCGC2011}.
However, while allelic score variants with known biological implications have been found to explain a considerable proportion of variance in specific traits, genome-wide scores demonstrated low specificity~\cite{Evans2013},
thus suggesting that there is currently little evidence which would recommend the use of genome-wide allelic scores as instruments.
Furthermore, allelic scores make use of raw genotype data thus limiting this approach to only a few locally available phenotypes.

In contrast, two-sample and meta-analysis approaches (see Chapter~\ref{cha:psychiatric_corr_mr}) can make use of summary statistics only to infer potential causal relationships. 
Since the effects of individual genetic variants are small, these methods commonly make use of multiple independent loci.
Here it is important to note that a strong effect size of a given SNP, used as instrumental variable, on the exposure does not suggest a strong causal connection between exposure and outcome, but rather represents the strength of the instrument.
A weak association between SNP and exposure, however, can result in an underpowered MR\@.

Similarly, it is unreasonable to assume that all genetic variants will fulfill all underlying assumptions.
Thus, multiple different methods have been developed in order to perform MR in the absence of clear biological information as well as to deal with possible assumption violations.

\subsubsection{Joint-Association studies}
\label{ssub:sensitivity_analysis}

\citet{Burgess2016a} suggested three different approaches to select genetic variants as potential instrumental  variables in the absence of biological knowledge.
(1) SNPs with indirect, but not conclusive, biological evidence.
These SNPs can be genome-wide associated genetic variants which have been replicated in multiple different populations.
(2) A more liberal approach would be to select genetic variants which are merely associated within a single population or sample.
This more liberal selection criteria is commonly used in the absence of clear biological information.
Sensitivity analysis can be used to investigate potential assumption violations, as done in Chapter~\ref{cha:psychiatric_corr_mr}.
(3) The most liberal approach is to use all available SNPs within the whole genome.
However, while this approach removes the necessity to justify  selection criteria of instrumental variables, it lacks theoretical legitimacy~\cite{Burgess2016a} and can result in a number of false findings~\cite{Evans2013}.

In Chapter~\ref{cha:psychiatric_corr_mr}, I made use of a liberal approach due to the lack of clear biological functional information.
On the basis of three GWAS on depression, schizophrenia, and bipolar disorder, I investigated potential causal effects on both risk taking and aggression.
Importantly, I chose a liberal $p$-value threshold for the inclusion of SNPs as instrumental variables.
A sensitivity analysis was used to investigate these findings for their robustness and, as shown in Figure~\ref{fig:sensitivity}, can provide valuable insight into possible assumption violations.

For example, funnel plots are commonly used to investigate potential directional pleiotropic effects.
Any asymmetry can be interpreted as a sign of directional pleiotropic effects.
However, the visual inspection of funnel plots can be questionable.
A few SNPs as instrumental variables can make the interpretation as either symmetric or asymmetric a difficult endeavour. 
This issue can also be seen in the funnel plots of Figure~\ref{fig:sensitivity}.
Specifically, the number of SNPs selected as instrumental variables for depressive symptoms and MDD are limited. 
Interestingly, one of the strongest assumptions of funnel plots, which are commonly used in a meta-analysis setting, might not be a problem in MR sensitivity analysis.

Within a traditional meta-analysis, funnel plots can provide an indication for potential publication bias.
However, they assume that effect size and sample size of each included study are uncorrelated~\cite{Evans2013}.
This is often not the case in many meta-analyses.
Indeed, studies might use different instruments to measure specific traits in small versus large samples or  might adjust sample size based on observed effect sizes of previous studies~\cite{Simonsohn}.
In contrast, the sample sizes of SNPs are rather constant.
Thus, funnel plots in MR do not suffer from the same problem.
Furthermore, in addition to funnel plots, potential pleiotropic effects can also be inspected by MR-egger's intercept (see Section~\ref{ssub:Used_Methods}).

An alternative explanation for an asymmetric funnel plot is that weaker instruments are more likely to be subject to the winner's curse~\cite{Taylor2014}.
The winner's curse was first described in the action theory literature~\cite{Bazerman1983} and states that the winning bid is likely to overestimate the true value of an item since it is the highest bid in the auction.
Similar in genome-wide association studies the focus is put commonly only  on significant SNPs, resulting in upwardly-biased effect size estimates.
This can lead to underpowered replication studies since sample size calculations are based on biased estimates~\cite{Xiao2008}.
Thus, instrument selection based on $p$-value alone can lead to an overestimation of the effect size between instrument and exposure, as well as  the association with the outcome~\cite{Bowden2015a,Taylor2014}.
\citet{Burgess2016c} showed that this can lead to a bias in MR estimates when overlapping samples were used to estimate genetic association with the outcome and exposure.
Indeed, overlapping samples were present when MR analysis was performed between depressive symptoms and impulsive aggression as well as risk taking,
thus potentially explaining the asymmetry in the funnel plots at Chapter~\ref{cha:psychiatric_corr_mr}.

Furthermore, these careful post-hoc analyses can be seen as too weak in contrast to the strong assumptions made my MR\@. 
Thus, \citet{Vanderweele2015} suggested that negative results might be more plausible when investigating potential causal connections.
In particular, in light of the strong assumptions as well as inability to fully verify those assumptions, empirically it might be more beneficial to test assumed causal relationships with the help of MR\@.
However, this would require well-powered studies with large sample sizes.
Indeed, the UK Biobank as well as recent other consortia have shown that it is possible to acquire large samples.
While null results are not immune to biases, such bias would need to align perfectly in order to achieve an effect size of zero with a narrow confidence interval when in fact a true effect is present.
An obvious drawback from this approach would be its inability to discover new interventions for various disorders. 

Furthermore, it is important to emphasize that MR cannot make conclusive statements about the causal relationship between two variables.
In order to establish a causal relationship, a string of multiple evidence is necessary.
MR can only play a part in this and numerous other additional studies are necessary to confirm a causal relationship.
Furthermore, it is important to remember that in contrast to randomized control trials, MR commonly aims to detect small effect sizes.

To conclude, `joint-analysis' studies are able to identify potential causal relationships without clear biological insight.
While empirical methods exist to test the underlying assumptions, only additional evidence by randomized control trials or other forms of studies will be able to fully confirm any potential causal relationship.

\section{Conclusion}
\label{sec:conclusion}

Aggressive behavior is a heritable trait which shows strong genetic correlations with a number of other phenotypes, such as smoking and risk taking.
However, it remains unclear  how to interpret genetic correlations in general and what implications these estimates have in explaining the underlying architecture of aggressive behavior.
Mendelian randomization might be able to elucidate possible causal relationships among traits, but application of such methods needs to be done with care while respecting the underlying strong assumptions. 
Nevertheless, a careful application of such techniques has shown that schizophrenia might have a causal effect on both risk taking and impulsive aggression.

My current genome-wide association studies lack statistical power to detect specific genetic variants associated with aggression, yet is the largest ever conducted.
Future studies should aim to increase sample size as well as improve phenotyping quality.
Furthermore, while simulations have shown that the newly-developed methods are better able to detect clusters of rare variants, those clusters seem not to be present for impulsive aggression.
