%&tex
\chapter{Introduction}
\label{cha:introduction}

% Dramatic entry
World War I and II, the Holocaust, as well as more recent terror attacks have demonstrated the extend of which human aggression is capable of causing distruction and harm to an extend previously unknown. 
Thus aggressive behavior remains a continous issue in all human societies and significant reseach effort has been spend to identify causes of human aggression.
This ranges from gun control, global warning, computer games and movies, as well as others to more biological and evolutionary causes. %TODO needs some citation for each
Here I will present major genetic characteristics of aggressive behavior in children and adults.
Ranging from studies on twins to large scale genome wide association studies.

% Intro to the intro
However, first I will present current and past research on aggressive behavior.
Within this literature review I will first present different subtypes of aggresssion.
Following I will describe commonly associated enviormental factors contributing to aggressive behavior.
This is succeeded by an introduction into the evolutionary theories of aggression.
I will then describe to which extend genetic factors influence aggressive behavior followed by a sumamry of recent discoveries in molecular genetics.
I will conclude this section by a short describtion of the here presented studies.

\section{Definition of Aggression}
\label{sec:overview_of_reseach_in_aggression}

One can define human aggression as a behavior intended to cause physical or emotional harm to others \cite{Anderson2002}.
This definition is relativly broad and does also include behavior in which the target does not intend to avoid the aggressive behavior, such as in sexual masochism.
In addition to the intend to cause harm the preptrator must also believe that the target wants to avoid the behavior\cite{Berkowitz1993,Baumeister1989,Baron2007,Geen2001} (e.i. to avoid being harmed).  
It is also important to further distinguish between \textit{Hostile Aggression} and \textit{Instrumental Aggresssion}.
While the former is characerised as impulsive, thoughtless and unplanned behavior, \textit{Instrumental Aggression} is defined as a planed and proactive behavior to obtain a certain higher goal \cite{Berkowitz1993,Geen2001}.
Both forms of aggression can be either physical or verbal.
While physical aggression in humans is homologous to other animals, verbal aggression, also called indirect, relational and social aggression, is relativly distinct to humans \cite{Archer2005}.
These verbal behaviors cause harm to others by gossiping, spreading rumors, or excluding other from social groups.
While the terms \textit{indirect}, \textit{relational}, and \textit{social} aggression have been differently conceptulized in the past \cite{Archer2001}, they are expressed in common behaviors, can be contrasted to physical aggression to a similar extend as well as have been shown to demonstrate the similar differences between the two sexes \cite{Archer2004}.
Hence the terms are more similar than distinct and I will therefore proceed to call all them \textit{indirect} aggression in order to distiguish it more from the physical, more direct aggressive behavior\cite{Archer2005}.

Thus I define aggression as an behavior by an individual who intends to either direct or indirect cause harm to other, while holding the believe that the target of such behavior aims to avoid such behavior.

Historically, research of aggression has been devided into nurture versus nature\cite{Archer2009}. 
Proponents of the nurture side have argued that aggressive behavior is caused by enviormental influences, while supports of the nature side of the discussion supported the idea that only differences in the genetic architecture are able to explain the differences in aggressive behavior.
Today's view is less polarized and acknoweghes that both, nature and nurture play a crucial part in aggression.
While not disregarding the enviormental aspect of aggression completly I will mostly focus on the genetic and biological causes of aggressive behavior.
In the following section I will outline evolutionary concepts helpful in explaining individual differences of aggression. 

\section{Evolutionary Theories on Aggressive Behavior}
\label{sec:evolutionary_theories_on_aggressive_behavior}



While aggression is most commonly used in an negative sense it has distinct evolutionary advatages.
For example, given limited available resources, aggressive behavior enables individuals to compete in a potential life threatening situations \cite[]{Anholt2012}.
In addition, hirarical order in social groups of individuals is often established by means of aggressive behavior which enables high ranked individuals prioriy access to food and mating paterns. 
Thus it has been suggested that aggressive behavior is under sexual selection\cite{Lindenfors2011}.

In fact the sexual selection theory (SST) has been used to great extend to explain the observed greater physical aggression in males in humans \cite{Archer2004}. 
Sexual selection is concerned how a member of one sex choses another individual from the other sex, as well as the competion between members of the sex over access to the other \cite{Darwin1859}.
In most mammals the more competitive sex is the male \cite{Archer2009} and~\cite{Trivers1972} suggested that these sex differences can be explained by the typical reduced parental investment of males.
Parental investment is the amount of resources a parent investes into his or her off spring to increase its survivial and reproduction \cite{Archer2009}.

In fact sex differences in aggressive behavior can be observed as early as at 2 years of age in parental ratings.
The exact origin of these early sex differences remain unknown, but effects of prenatal testosteron and eviormental factors have been proposed.
However, one cannot exclude that sex differences at this early age are not caused by rater bias. %TODO cite 
Thus, parents might perceive male toddler as more aggressive since they would expect more aggressive behavior from male children.

Nevertheless, one of the most crucial differences between male and female is the extend to which an individual is willing to take during a conflict %TODO cite.



Despite these advatages, increased levels of aggressive behavior can also be harmful to the individual.
Expressing aggression towards another individual sheers energy away from other acitivies, such as hunting and foraging, and also carries a high risk of injury and death \cite{Packer1995}.  
Hence previous reseach on multiple different species suggested that aggressive behavior is under stabilizing selection.
Further this also shows strong evidence for a large role of genetic factors in the etiology of aggressive behavior. 


In addition one can also see aggressive behavior as part of a wider phenotypical complex. 


